---
title: "Homework 4. Part 2"
subtitle: "Advanced Network Analytics and Modeling"
author: "Aleksey Shadrunov, Anastasia Podolskaya, Maksim Emelyanenko"
date: "`r Sys.Date()`"
indent: true
output: 
  pdf_document:
    toc: true
    toc_depth: 3
    extra_dependencies: ["float"]
header-includes:
    - \usepackage{indentfirst}
---

\newpage

# Prepare packages and load data

Install packages (uncomment if necessary):
```{r}
# install.packages("network", repos = "https://mirror.truenetwork.ru/CRAN/")
# install.packages("sna", repos = "https://mirror.truenetwork.ru/CRAN/")
# install.packages("igraph", repos = "https://mirror.truenetwork.ru/CRAN/")
# install.packages("intergraph", repos = "https://mirror.truenetwork.ru/CRAN/")
# install.packages("knitr", repos = "https://mirror.truenetwork.ru/CRAN/")
# install.packages("RColorBrewer", repos = "https://mirror.truenetwork.ru/CRAN/")
# install.packages("bipartite", repos = "https://mirror.truenetwork.ru/CRAN/")
# install.packages("dendextend", repos = "https://mirror.truenetwork.ru/CRAN/")
# install.packages("blockmodeling", repos = "https://mirror.truenetwork.ru/CRAN/")
# install.packages("multinet", repos = "https://mirror.truenetwork.ru/CRAN/")
```


```{r}
suppressPackageStartupMessages(library(RColorBrewer))
suppressPackageStartupMessages(library(bipartite))
suppressPackageStartupMessages(library(network))
suppressPackageStartupMessages(library(dendextend))
suppressPackageStartupMessages(library(knitr))
suppressPackageStartupMessages(library(stringr))
```

The given dataset is in CSV format, which is a table with columns: Entity.A, Entity.B, Entity.A.Type, Entity.B.Type, Connection. As we have already looked through the net, we can clean up the data a bit.
```{r}
Trump <- read.csv("/home/alex/Documents/year-3-sna/hw4/trumpworld.csv")

ncol(Trump)
nrow(Trump)
Trump <- Trump[-c(6)]

Trump$Connection <- Trump$Connection %>% str_replace(" \\(as of 2016 FEC filing\\)", "")
Trump$Connection <- Trump$Connection %>% str_replace(" \\(as of 2014 FEC filing\\)", "")
Trump$Connection <- Trump$Connection %>% str_replace("Former director", "Director")
Trump$Connection <- Trump$Connection %>% str_replace("Former member", "Member")
Trump$Connection <- Trump$Connection %>% str_replace("Reported member", "Member")
Trump$Connection <- Trump$Connection %>% str_replace("Special assistant", "Assistant")
Trump$Connection <- Trump$Connection %>% str_replace("Deputy assistant", "Assistant")
Trump$Connection <- Trump$Connection %>% str_trim()
```

# Create networks
Next, we want to select several Connection types to create a multiplex networks. Here we see the most common connections:
```{r}
cons <- data.frame(table(Trump$Connection))
kable(head(cons[order(cons$Freq, decreasing = TRUE), ], 25))
```

Also we may select rows in each type of entities.

## Organization to Organization
```{r}
b2b <- Trump[Trump$Entity.A.Type == "Organization", ]
b2b <- b2b[b2b$Entity.B.Type == "Organization", ]

b2b_cons <- data.frame(table(b2b$Connection))
kable(head(b2b_cons[order(b2b_cons$Freq, decreasing = TRUE), ], 10))
```

We may select top 2 for further analysis.

```{r}
b2b_top_count <- 2
b2b_cons_selected <- as.vector(head(b2b_cons[order(b2b_cons$Freq, decreasing = TRUE), ], b2b_top_count)$Var1)
b2b_cons_selected

b2b_selected <- subset(b2b, Connection %in% b2b_cons_selected)

b2b_A_selected <- as.vector(unique(b2b_selected$Entity.A))
b2b_B_selected <- as.vector(unique(b2b_selected$Entity.B))
```

## Person to Person
```{r}
c2c <- Trump[Trump$Entity.A.Type == "Person", ]
c2c <- c2c[c2c$Entity.B.Type == "Person", ]

c2c_cons <- data.frame(table(c2c$Connection))
kable(head(c2c_cons[order(c2c_cons$Freq, decreasing = TRUE), ], 10))
```

We may select top 4 for further analysis.

```{r}
c2c_top_count <- 4
c2c_cons_selected <- as.vector(head(c2c_cons[order(c2c_cons$Freq, decreasing = TRUE), ], c2c_top_count)$Var1)
c2c_cons_selected

c2c_selected <- subset(c2c, Connection %in% c2c_cons_selected)

c2c_A_selected <- as.vector(unique(c2c_selected$Entity.A))
c2c_B_selected <- as.vector(unique(c2c_selected$Entity.B))
```

## Person or Organization
```{r}
b2c <- Trump[Trump$Entity.A.Type == "Organization", ]
b2c <- b2c[b2c$Entity.B.Type == "Person", ]
c2b <- Trump[Trump$Entity.A.Type == "Person", ]
c2b <- c2b[c2b$Entity.B.Type == "Organization", ]
mix <- rbind(b2c, c2b)

mix_cons <- data.frame(table(mix$Connection))
kable(head(mix_cons[order(mix_cons$Freq, decreasing = TRUE), ], 10))
```

We may select top 3 for further analysis.

```{r}
mix_top_count <- 3
mix_cons_selected <- as.vector(head(mix_cons[order(mix_cons$Freq, decreasing = TRUE), ], mix_top_count)$Var1)
mix_cons_selected

mix_selected <- subset(mix, Connection %in% mix_cons_selected)

mix_A_selected <- as.vector(unique(mix_selected$Entity.A))
mix_B_selected <- as.vector(unique(mix_selected$Entity.B))
```

Now we can combine all subsets together:
```{r}
all_selected <- rbind(b2b_selected, c2c_selected, mix_selected)
all_actors_selected <- unique(c(all_selected$Entity.A, all_selected$Entity.B))
```

For creating 3-mode network (Person, Organization, Federal Agency), we should add later rows to the dataframe. Alse we may filter only those actors who are included in the rest of the network.
```{r}
fa <- Trump[Trump$Entity.B.Type == "Federal Agency", ]
fa_selected <- fa[fa$Entity.A %in% all_actors_selected, ]
```

# Multiplex network analysis

Now we may create multiplex network for each type of actors: Persons or Organization.

## Organization to Organization
### Plot

```{r}
suppressPackageStartupMessages(library(multinet))

b2bm <- ml_empty()
# add two directed layers
add_layers_ml(b2bm, b2b_cons_selected, rep(c(TRUE), length(b2b_cons_selected)))
is_directed_ml(b2bm)

# add actors (repeat each actor on each layer)
b2b_actors_selected <- unique(c(b2b_A_selected, b2b_B_selected))
actors <- rep(b2b_actors_selected, length(b2b_cons_selected))
layers <- c()
for (con in b2b_cons_selected) {
    layers <- c(layers, rep(con, length(b2b_actors_selected)))
}
vertices <- data.frame(actors = actors, layers = layers)
add_vertices_ml(b2bm, vertices)

# add arcs
actors_from <- c()
layers_from <- c()
actors_to <- c()
layers_to <- c()
b2b_selected_for_layer <- c()

for (layer in b2b_cons_selected) {
    b2b_selected_for_layer <- b2b_selected[b2b_selected$Connection == layer, ]
    actors_from <- c(actors_from, b2b_selected_for_layer$Entity.A)
    actors_to <- c(actors_to, b2b_selected_for_layer$Entity.B)
    layers_from <- c(layers_from, b2b_selected_for_layer$Connection)
    layers_to <- c(layers_to, b2b_selected_for_layer$Connection)
}

edges <- data.frame(
    actors_from = actors_from,
    layers_from = layers_from,
    actors_to = actors_to,
    layers_to = layers_to
)

add_edges_ml(b2bm, edges)
```

Now we can plot the two-layer network to take a first glance at its properties (Figure 1). 
```{r fig.cap = "b2b Visualisation", fig.height = 9.5, fig.width = 7, fig.pos = "!h"}
suppressWarnings({
    par(mar = c(0, 0, 0, 0))
    l <- layout_multiforce_ml(b2bm, w_inter = 100, gravity = 100, iter = 100)
    plot(b2bm,
        layout = l,
        grid = c(2, 1),
        vertex.labels = "",
        vertex.labels.cex = 0.3, vertex.labels.col = "grey",
        vertex.color = "orange",
        vertex.cex = 0.5, edge.col = "grey",
    )
})
```

What we see is a very unconcise picture of the network. As both layers are little interconnected, we may plot the separately. 

```{r}
layers <- as.list(b2bm)
names(layers)
```

```{r fig.cap = "b2b Separate Visualisation", fig.height = 11, fig.width = 9, fig.pos = "!h"}
opar <- par(mfrow = c(2, 1))
par(mar = c(0, 0, 0, 0))

plot(
    layers[[2]],
    vertex.size = 6,
    edge.arrow.size = 0.2,
    edge.color = "grey",
    vertex.label.cex = 0.3,
    vertex.color = "orange",
    vertex.shape = "circle",
    vertex.frame.color = "grey",
    vertex.label.dist = 0,
    vertex.label.color = "blue"
)
plot(
    layers[[3]],
    vertex.size = 4,
    edge.arrow.size = 0.2,
    edge.color = "grey",
    vertex.label.cex = 0.2,
    vertex.color = "green",
    vertex.shape = "circle",
    vertex.frame.color = "grey",
    vertex.label.dist = 0,
    vertex.label.color = "blue"
)
```

Now the picture is clearer and we can observe, that in both layers there are no isolated vertices. The first layer (Investor) presents one central actor which is apparently a huge investing company (the name is THRIVE CAPITAL). The second layer (Ownership), in contrast, have many connected components, which are separated from each other. The most central vertices (with highest degree) are DJT HOLDINGS LLC, IT OPERATIONS MANAGING MEMBER CORP. So, we may suggest, that it is more common when one company invests in many others and not when one company owns many others.

### Statistics
Simple summary command provides us with the following Statistics:
* n - number of vertices
* m - number of edges
* dir - directionality
* nc - number of connected components (subgraphs in which any two vertices are connected to each other by paths. In case of directed layers, edge directionality
will be taken into account, i.e. the number of strong components)
* dens - density (the ratio between the number of edges and the number of possible edges)
* cc - clustering coefficient (the ratio between the triangles and the connected triples in the layer)
* apl - average path length (the average graph-distance between all pairs of vertices in the layer)
* dia - diameter (the longest graph distance between any two vertices in the layer)
```{r}
summary(b2bm)
```
So, we see that the first layer has only 3 connected components, while the second — 170. However, the density of Investor is 10 times more than the density of Ownerships (much more ties could theoretically be created in the second layer). Curious value are assigned to the longest distance in two layers. The second one has the distance of 5, which means, that there is an ownership chain consisted of 5 companies.

#### Degree

We continue with analysing degree for each layer. 
```{r}
color_map <- brewer.pal(num_layers_ml(b2bm), "Set3")
opar <- par(mfrow = c(1, 3))
hist(degree_ml(b2bm),
    breaks = max(degree_ml(b2bm)),
    main = "flattened",
    xlab = "degree"
)
for (i in 1:num_layers_ml(b2bm)) {
    d <- degree_ml(b2bm, layers = layers_ml(b2bm)[[i]])
    hist(d,
        breaks = max(d, na.rm = TRUE), main = layers_ml(b2bm)[[i]],
        xlab = "degree", col = color_map[i]
    )
}
```
```{r}
# flatten
kable(t(table(degree(layers[[1]]))), caption = "Flatten: Degree distribution")

# Investor
kable(t(table(degree(layers[[2]]))), caption = "Investor: Degree distribution")

# Ownership
kable(t(table(degree(layers[[3]]))), caption = "Ownership: Degree distribution")
```

We have already observed that most actors are of degree 1, and few possess an outstanding value (84 or 34).
Let us look at those vertices:

```{r}
# Investor
d <- data.frame(degree(layers[[2]]))
d <- cbind(newColName = rownames(d), d)
colnames(d) <- c("Actor", "Value")
rownames(d) <- 1:nrow(d)
kable(head(d[order(d$Value, decreasing = TRUE), ], 5), caption = "Investor: Degree top")

# Ownership
d <- data.frame(degree(layers[[3]]))
d <- cbind(newColName = rownames(d), d)
colnames(d) <- c("Actor", "Value")
rownames(d) <- 1:nrow(d)
kable(head(d[order(d$Value, decreasing = TRUE), ], 5), caption = "Ownership: Degree top")
```

#### Layer comparison
We may also compare layers. However, in our case, they are not very related, so most of statistics are not useful.

We start with Pearson correlation coefficient. The smallest value (-1) indicates that high-degree actors in one layer are low-degree in the other, while the largest value (1) is returned if high-degree actors in one layer are high-degree actors in the other. 
```{r}
layer_comparison_ml(b2bm, method = "pearson.degree")
```
We see that correlation is about zero, which is expected because the actors in both layers are different. 


#### Dyad
For each layer we may calculate dyad statistics:

```{r}
Investor <- as.igraph(b2bm, "Investor")
Ownership <- as.igraph(b2bm, "Ownership")

# types of dyads
dyad.census(Investor)
dyad.census(Ownership)
```
Dyad census says that no mutual ties are in the Ivestor layer. Ownership layer behaves similarly, but has more arcs.

#### Triads
```{r}
Triads <- c("003", "012", "102", "021D", "021U", "021C", "111D", "111U", "030T", "030C", "201", "120D", "120U", "120C", "210", "300")
Number <- triad.census(Investor)
kable(t(data.frame(Triads, Number)), caption = "Investor: triads")

Number <- triad.census(Ownership)
kable(t(data.frame(Triads, Number)), caption = "Ownership: triads")
```
In Investor layer, there are plenty of A<-B->C (outcoming) triads, some A->B<-C triads, also there are transitive triads A->B->C. Ownership layer is similar, too. 


#### Transitivity
Let's calculate trads statics:
```{r}
transitivity(Investor)
transitivity(Ownership)
```
In both layers transitivity is about zero, as there are not many connected ties. 

#### Reciprocity
```{r}
reciprocity(Investor)
reciprocity(Ownership)
```
In both layers reciprocity is zero^ there are no mutual arcs. 

#### Centralitiy

It makes sense to find the most central actors for each type of centrality. Probably the most-ranked vertices will be the same.
```{r}
# Investor
d <- data.frame(degree(Investor, mode = "in"))
d <- cbind(newColName = rownames(d), d)
colnames(d) <- c("Actor", "Value")
rownames(d) <- 1:nrow(d)
kable(head(d[order(d$Value, decreasing = TRUE), ], 5), caption = "Investor: Indegree centrality")

d <- data.frame(degree(Investor, mode = "out"))
d <- cbind(newColName = rownames(d), d)
colnames(d) <- c("Actor", "Value")
rownames(d) <- 1:nrow(d)
kable(head(d[order(d$Value, decreasing = TRUE), ], 5), caption = "Investor: Outdegree centrality")

d <- data.frame(degree(Investor, mode = "total"))
d <- cbind(newColName = rownames(d), d)
colnames(d) <- c("Actor", "Value")
rownames(d) <- 1:nrow(d)
kable(head(d[order(d$Value, decreasing = TRUE), ], 5), caption = "Investor: Total degree centrality")

d <- data.frame(closeness(Investor))
d <- cbind(newColName = rownames(d), d)
colnames(d) <- c("Actor", "Value")
rownames(d) <- 1:nrow(d)
kable(head(d[order(d$Value, decreasing = TRUE), ], 5), caption = "Investor: Closeness centrality")
```

```{r}
# Ownership
d <- data.frame(degree(Ownership, mode = "in"))
d <- cbind(newColName = rownames(d), d)
colnames(d) <- c("Actor", "Value")
rownames(d) <- 1:nrow(d)
kable(head(d[order(d$Value, decreasing = TRUE), ], 7), caption = "Ownership: Indegree centrality")

d <- data.frame(degree(Ownership, mode = "out"))
d <- cbind(newColName = rownames(d), d)
colnames(d) <- c("Actor", "Value")
rownames(d) <- 1:nrow(d)
kable(head(d[order(d$Value, decreasing = TRUE), ], 5), caption = "Ownership: Outdegree centrality")

d <- data.frame(degree(Ownership, mode = "total"))
d <- cbind(newColName = rownames(d), d)
colnames(d) <- c("Actor", "Value")
rownames(d) <- 1:nrow(d)
kable(head(d[order(d$Value, decreasing = TRUE), ], 5), caption = "Ownership: Total degree centrality")

d <- data.frame(closeness(Ownership))
d <- cbind(newColName = rownames(d), d)
colnames(d) <- c("Actor", "Value")
rownames(d) <- 1:nrow(d)
kable(head(d[order(d$Value, decreasing = TRUE), ], 5), caption = "Ownership: Closeness centrality")
```

We may notice some centralities are consistent with previous top positions, and some are not.

#### Centralization
```{r}
centralization.degree(Investor)$centralization
centralization.degree(Ownership)$centralization
```
The graph level centrality index is the last metric to look at. It demonstrates the measure of centrality of the whole graph, and in our case the Investor layer has a greater value. It is proven by visualisations we made previously.

### Summary

The first multiplex network from Trumpworld data consists of two layers. Each of them has different structure: Investor layer is very simple with 3 components, one of them comprises almost all vertices. It is a very central actor which is connected to many other, however, only in one direction. The second layer has many small components and several comparatively large, which is built similarly to the one from Investor layer. In this component we have found a path of length 5. However, such structures are less common for layer 2.




## Person to Person
### Plot

```{r}
c2cm <- ml_empty()
# add four directed layers
add_layers_ml(c2cm, c2c_cons_selected, rep(c(TRUE), length(c2c_cons_selected)))
is_directed_ml(c2cm)

# add actors (repeat each actor on each layer)
c2c_actors_selected <- unique(c(c2c_A_selected, c2c_B_selected))
actors <- rep(c2c_actors_selected, length(c2c_cons_selected))
layers <- c()
for (con in c2c_cons_selected) {
    layers <- c(layers, rep(con, length(c2c_actors_selected)))
}
vertices <- data.frame(actors = actors, layers = layers)
add_vertices_ml(c2cm, vertices)

# add arcs
actors_from <- c()
layers_from <- c()
actors_to <- c()
layers_to <- c()
c2c_selected_for_layer <- c()

for (layer in c2c_cons_selected) {
    c2c_selected_for_layer <- c2c_selected[c2c_selected$Connection == layer, ]
    actors_from <- c(actors_from, c2c_selected_for_layer$Entity.A)
    actors_to <- c(actors_to, c2c_selected_for_layer$Entity.B)
    layers_from <- c(layers_from, c2c_selected_for_layer$Connection)
    layers_to <- c(layers_to, c2c_selected_for_layer$Connection)
}

edges <- data.frame(
    actors_from = actors_from,
    layers_from = layers_from,
    actors_to = actors_to,
    layers_to = layers_to
)

add_edges_ml(c2cm, edges)
```

Now we can plot the 4-layer network to take a first glance at its properties (Figure 3). 
```{r fig.cap = "c2c Visualisation", fig.height = 9.5, fig.width = 7, fig.pos = "!h"}
suppressWarnings({
    par(mar = c(0, 0, 0, 0))
    l <- layout_multiforce_ml(c2cm, w_inter = 100, gravity = 100, iter = 100)
    plot(c2cm,
        layout = l,
        grid = c(2, 2),
        vertex.labels = "",
        vertex.labels.cex = 0.3, vertex.labels.col = "grey",
        vertex.color = "purple",
        vertex.cex = 0.5, edge.col = "grey",
    )
})
```

On contrary with previous network, we see different pictures here. The most comprehensive is the Assistant layer. Here we may tell that there is again a very central actor which connects to many other in one-way manner.  

To make picture clearer, we should perform separate visualisations.
```{r}
layers <- as.list(c2cm)
names(layers)
```

```{r fig.cap = "c2c Separate Visualisation", fig.height = 9, fig.width = 7, fig.pos = "!h"}
opar <- par(mfrow = c(2, 2))
par(mar = c(0, 0, 0, 0))

plot(
    layers[[2]],
    vertex.size = 8,
    edge.arrow.size = 0.3,
    edge.color = "grey",
    vertex.label.cex = 0.3,
    vertex.color = "orange",
    vertex.shape = "circle",
    vertex.frame.color = "grey",
    vertex.label.dist = 0,
    vertex.label.color = "blue"
)
plot(
    layers[[3]],
    vertex.size = 8,
    edge.arrow.size = 0.3,
    edge.color = "grey",
    vertex.label.cex = 0.3,
    vertex.color = "green",
    vertex.shape = "circle",
    vertex.frame.color = "grey",
    vertex.label.dist = 0,
    vertex.label.color = "blue"
)
plot(
    layers[[4]],
    vertex.size = 8,
    edge.arrow.size = 0.3,
    edge.color = "grey",
    vertex.label.cex = 0.3,
    vertex.color = "magenta",
    vertex.shape = "circle",
    vertex.frame.color = "grey",
    vertex.label.dist = 0,
    vertex.label.color = "blue"
)
plot(
    layers[[5]],
    vertex.size = 8,
    edge.arrow.size = 0.3,
    edge.color = "grey",
    vertex.label.cex = 0.3,
    vertex.color = "red",
    vertex.shape = "circle",
    vertex.frame.color = "grey",
    vertex.label.dist = 0,
    vertex.label.color = "blue"
)
```

Now the picture is clearer (Figure 4). 

The first graph (Assistant) is a very central one, with one actor (DONALD J. TRUMP) being in the centre. The single other component consists of two vertices, not linked to main subgraph (JARED KUSHNER, AVRAHM BERKOWITZ).  

The second one (Married) consist only from pairs of married people. Here we observe only not reciprocal arcs. Due to its nature, the network is very unconnected, as there is no path longer 2 and no vertex with degree more than 1. So, no one had a divorce according to this picture.

The third one (Parent/child) consists of parents and their children. Here we find various figures, even consisting of 4 actors.

The fourth one (Siblings) is an interesting one. Here we have highly coupled graphs, almost a clique, and also triads. 

### Statistics
Simple summary gives us the following:
```{r}
summary(c2cm)
```
So, we see that the Assistant layer has 2 connected components, while the others have more — from 9 to 17. All layers have more or less similar density (0.015-0.036). The average path length and diameter are not significant in every layer.

#### Degree  

We continue with analysing degree for each layer. 
```{r}
color_map <- brewer.pal(num_layers_ml(c2cm), "Set3")
opar <- par(mfrow = c(2, 2))
for (i in 1:num_layers_ml(c2cm)) {
    d <- degree_ml(c2cm, layers = layers_ml(c2cm)[[i]])
    hist(d,
        breaks = max(d, na.rm = TRUE), main = layers_ml(c2cm)[[i]],
        xlab = "degree", col = color_map[i]
    )
}
```
```{r}
kable(t(table(degree(layers[[1]]))), caption = "Flatten: Degree distribution")
kable(t(table(degree(layers[[2]]))), caption = "Assistant: Degree distribution")
kable(t(table(degree(layers[[3]]))), caption = "Married: Degree distribution")
kable(t(table(degree(layers[[4]]))), caption = "Parent/child: Degree distribution")
kable(t(table(degree(layers[[5]]))), caption = "Siblings: Degree distribution")
```


Here all values on histograms are consistent with our observations from the graphs. Parents have from 1 to 4 children, Assistants of DONALD J. TRUMP have degree of 1, Married people are connected only to one person. Siblings have from 1 to 3 brothers or sisters. 
Let us look at the distribution in tables:

```{r}
# Investor
d <- data.frame(degree(layers[[2]]))
d <- cbind(newColName = rownames(d), d)
colnames(d) <- c("Actor", "Value")
rownames(d) <- 1:nrow(d)
kable(head(d[order(d$Value, decreasing = TRUE), ], 5), caption = "Investor: Degree top")

# Ownership
d <- data.frame(degree(layers[[3]]))
d <- cbind(newColName = rownames(d), d)
colnames(d) <- c("Actor", "Value")
rownames(d) <- 1:nrow(d)
kable(head(d[order(d$Value, decreasing = TRUE), ], 5), caption = "Ownership: Degree top")
```
